---
title: "Exocarpochory"
author: "David S, Mason"
date: "6/14/2021"
output:
  pdf_document: default
  html_document: default
---
## Generating one set of dispersal observations from two different dispersal kernels

This simulation explores the impact of exocarpochory on dispersal distance. We use probability distributions that reflect wind dispersal and epizoochory to draw random observatons. These draws are repeated 10000 times with differing probability of exocarpochory occuring. Then, we calculate the difference between observations with and without exocarpochory. Next,we depict the positive relationship between increase in disperal distance and probability for exocarpochory. Finally, we generate histograms of these observations for direct
comparison of the simulations.

## Bring in the packages
```{r}
library(LaplacesDemon)
library(tidyverse)
options(scipen = 999) # No scientific notation
```
## Basic (non-mixed) kernels
```{r}
# Conyza
set.seed(14356)

wind <- as.data.frame(rcauchy(n=10000,
															location=0.00000249,scale=2.7658))
colnames(wind) <- c("Distance")

wind <- wind %>% filter(Distance>=0)

wind$Kernel <- "Wind only"
LDD.wind <- filter(wind, Distance>=100)
length(LDD.wind)
# Trillium 	
animal <- as.data.frame(rpe(n=10000,sigma=188.36,kappa=0.7341))
colnames(animal) <- c("Distance")

animal <- animal %>% filter(Distance>0)
```
## Set parameters for simulation
```{r}
# Create vectors of the probability of exocarpochory
p.wind <- c(0.99, 0.95, 0.90, 0.80, 0.70)
p.animal <- 1-p.wind

# Set the number of draws
n <- 10000
```

## Simulation with 1% Exocarpochory
```{r}
mix.sims.out <- rep(0,n) # Create output vector
set.seed(4) # We set a new seed everytime because.....

for(i in 1:n){ # For loop runs n draws
  
  U <- runif(n=1) # Select random number between 0-1
  if(U <= p.wind[1]){ # If the probability of wind is less than that number
    # Pick Conyza (wind) kernel:
    mix.sims.out[i] <- rcauchy(n=1,
															location=0.00000249,scale=2.7658)
  }else{ # Otherwise
    # Pick Desmodium (epizoo) animal kernel: 
    mix.sims.out[i] <- rpe(n=1,sigma=188.36,kappa=0.7341)
  }
}

mix.sims.out.1 <- as.data.frame(mix.sims.out) # Save as unique dataframe
colnames(mix.sims.out.1) <- c("Distance") # Change column name
mix.sims.out.1$Kernel <- "1% Exocarpochory" # Create column to ID kernel
mix.sims.out.1 <- mix.sims.out.1 %>% filter(Distance>=0) # Truncate the data

# Calculate comparison metrics
median(mix.sims.out.1$Distance)-median(wind$Distance) # Difference in median = 0.05361729
quantile(mix.sims.out.1$Distance, probs=c(.95))-
	quantile(wind$Distance, probs=c(.95)) # Difference in 95% quantile = 12.13406 
LDD.1 <- filter(mix.sims.out.1, Distance>=100) # How many observations > 100 m
length(LDD.1)
```


## Simulation with 5% Exocarpochory
```{r}
mix.sims.out <- rep(0,n) # Make a new vector to write over each simulation
set.seed(302)

for(i in 1:n){
  
  U <- runif(n=1)
  if(U <= p.wind[2]){ # The indexing here is how the probability changes
  	# Pick Conyza (wind) kernel:
    mix.sims.out[i] <- rcauchy(n=1,
															location=0.00000249,scale=2.7658)
  }else{
    # Pick Desmodium (epizoo) animal kernel: 
    mix.sims.out[i] <- rpe(n=1,sigma=188.36,kappa=0.7341)
  }
}

mix.sims.out.5 <- as.data.frame(mix.sims.out)
colnames(mix.sims.out.5) <- c("Distance")
mix.sims.out.5$Kernel <- "5% Exocarpochory"
mix.sims.out.5 <- mix.sims.out.5 %>% filter(Distance>=0)

median(mix.sims.out.5$Distance)-median(wind$Distance) # 0.287181
quantile(mix.sims.out.5$Distance, probs=c(.95))-
	quantile(wind$Distance, probs=c(.95)) # 52.16673  
LDD.5 <- filter(mix.sims.out.5, Distance>=100)
length(LDD.5)
```

## Simulation with 10% Exocarpochory
```{r}
mix.sims.out <- rep(0,n)
set.seed(49)
for(i in 1:n){
  
  U <- runif(n=1)
  if(U <= p.wind[3]){
    # Pick wind kernel:
    mix.sims.out[i] <- rcauchy(n=1,
															location=0.00000249,scale=2.7658)
  }else{
    # Pick animal kernel: 
    mix.sims.out[i] <- rpe(n=1,sigma=188.36,kappa=0.7341)
  }
}

mix.sims.out.10 <- as.data.frame(mix.sims.out)
colnames(mix.sims.out.10) <- c("Distance")
mix.sims.out.10$Kernel <- "10% Exocarpochory"
mix.sims.out.10 <- mix.sims.out.10 %>% filter(Distance>=0)

median(mix.sims.out.10$Distance)-median(wind$Distance) #  0.5404275
quantile(mix.sims.out.10$Distance, probs=c(.95))-
	quantile(wind$Distance, probs=c(0.95)) # 150.6252   
LDD.10 <- filter(mix.sims.out.10, Distance>=100)
length(LDD.10)
```
## Simulation with 20% Exocarpochory
```{r}
mix.sims.out <- rep(0,n)
set.seed(501)
for(i in 1:n){
  
  U <- runif(n=1)
  if(U <= p.wind[4]){
    #pick wind kernel:
    mix.sims.out[i] <- rcauchy(n=1,
															location=0.00000249,scale=2.7658)
  }else{
    #pick animal kernel: 
    mix.sims.out[i] <- rpe(n=1,sigma=188.36,kappa=0.7341)
  }
}

mix.sims.out.20 <- as.data.frame(mix.sims.out)
colnames(mix.sims.out.20) <- c("Distance")
mix.sims.out.20$Kernel <- "20% Exocarpochory"
mix.sims.out.20 <- mix.sims.out.20 %>% filter(Distance>=0)

median(mix.sims.out.20$Distance)-median(wind$Distance) # 1.337396
quantile(mix.sims.out.20$Distance, probs=c(.95))-
	quantile(wind$Distance, probs=c(.95)) # 272.4912 
LDD.20 <- filter(mix.sims.out.20, Distance>=100)
length(LDD.20)
```
## Simulation with 30% Exocarpochory
```{r}
mix.sims.out <- rep(0,n)
set.seed(000109)
for(i in 1:n){
  
  U <- runif(n=1)
  if(U <= p.wind[5]){
    # Pick wind kernel:
    mix.sims.out[i] <- rcauchy(n=1,
															location=0.00000249,scale=2.7658)
  }else{
    # Pick animal kernel: 
    mix.sims.out[i] <- rpe(n=1,sigma=188.36,kappa=0.7341)
  }
}

mix.sims.out.30 <- as.data.frame(mix.sims.out)
colnames(mix.sims.out.30) <- c("Distance")
mix.sims.out.30$Kernel <- "30% Exocarpochory"
mix.sims.out.30 <- mix.sims.out.30 %>% filter(Distance>=0)

median(mix.sims.out.30$Distance)-median(wind$Distance) # 2.5146
quantile(mix.sims.out.30$Distance, probs=c(.95))-
	quantile(wind$Distance, probs=c(.95)) # 405.3121 
LDD.30 <- filter(mix.sims.out.30, Distance>=100)
length(LDD.30)
```
## Collate comparison metrics 
```{r}
# Create a data frame with the change in median and 95% quantile
df <- data.frame(Probability = c(1,5,10,20,30),
								 Delta_median = c(0.05361729,0.287181,0.5404275, 1.337396,2.5146),
								 Delta_95 = c(12.13406,52.16673,150.6252, 272.4912,405.3121))

# Set probability as a factor
df$Probability <- as.factor(df$Probability) 

# Convert the data to long format
df_long <- pivot_longer(df, cols = c("Delta_median",
																		 "Delta_95"), names_to = "Parameter")

# Load plyr to revalue parameters
library(plyr)
df_long$Parameter <- plyr::revalue(df_long$Parameter, c("Delta_median"="50th percentile", 
												 "Delta_95"="95th percentile"))

# Assign different limits for each parameter to manipulate figure axis
limits <- data.frame(Parameter = c("50th percentile", "95th percentile"),
																	 y_min = c(0,0), y_max = c(3,450))

# Add them to the original dataframe
fifty <- filter(df_long, Parameter=='50th percentile')
fifty$y_min <- 0
fifty$y_max <- 3
ninety <- filter(df_long, Parameter=='95th percentile')
ninety$y_min <- 0
ninety$y_max <- 450

df_long <- rbind(fifty,ninety)
```

## Plot comparison metrics 
```{r}
p1<-ggplot(df_long, aes(x = Probability, y = value))+
			 		geom_point(size=10)+
					ylab("Increase in dispersal distance (m)")+
					theme_classic()+
					xlab("Probability of Exocarpochory (%)")+
					theme(text = element_text(size = 25))+
					theme(axis.title.y = element_text(vjust = 2))+
					scale_x_discrete(breaks=c("1","5","10", "20", "30"))+
					facet_wrap(~Parameter, scales = "free_y", nrow = 2)+
					geom_blank(aes(y = y_min))+ # Alter axis limits for facet 
				  geom_blank(aes(y = y_max))  # wrapped parameters
```
## Plot observations from simulations

RE: Warning messages. Extreme values (i.e., beyond the axis represented in the figure) are dropped. which I believe is triggering the warning messages. Unfortunately, we are forced to strike a balance between practical/effective visualization and thoroughness/honest representation. If we leave in extreme values, the x-axis will be long enough to accomodate values > 20000 m, which inhibits interpretation
of the rest of the data. 

*This will not add the long-distance dispersal information seen on the histograms in the final version. This information was addded on top of the combined figure using Preview.*
```{r}

# Combine simulations into vector, drop 
comb.kernel.no.wind <- rbind(mix.sims.out.1,mix.sims.out.10,mix.sims.out.30)

# Make a vector of wind simulation for histogram background
wind.fig <- as.data.frame(wind[,1])
colnames(wind.fig) <- c("Distance")

# Get values to backtransform labels on x-axis
log(11) # 2.397895
log(21) # 3.044522
log(751) # 6.621406


library(RColorBrewer)
brewer.pal(3, "Blues") 

p2<-ggplot(comb.kernel.no.wind, aes(x=log(1+Distance),fill=Kernel))+
	geom_histogram(wind.fig, mapping = aes(x=log(1+Distance)), fill='lightgray',
								 bins = 80, alpha = 0.8, color = 'black')+
	geom_histogram(position="identity",
  							 bins = 80, color = "black", alpha = 0.7)+
	geom_vline(xintercept=4.615121, linetype='solid', color='black', size=1)+
	scale_fill_manual(values = c('#DEEBF7','#9ECAE1','#3182BD'))+
	scale_x_continuous(expand = c(0,0),
										 breaks = c(0,2.397895,
																3.044522,6.621406),
										 labels = c("0","10","100",
										 					 "750"),
										 limits = c(0, 7.5))+
	scale_y_continuous(expand = c(0,0),limits = c(0,250))+
	facet_wrap(~Kernel,ncol = 1)+
	xlab("Distance (m)")+
	ylab("Frequency")+
	theme_classic()+
	theme(strip.background = element_blank(),strip.text.x = element_blank(),
        panel.background = element_rect(colour = "black", size=1.5),
				text = element_text(size=25),panel.spacing = unit(1.5, "lines"),
				legend.position="top", legend.title = element_blank())
```
## Combine plots
```{r}
library(ggpubr)
combo <- ggarrange(p2, p1,labels = c("a", "b"),
					common.legend = TRUE, legend='top')

ggsave('combo.jpg', width = 14, height = 8, dpi = 300)
```